Is Partner Modeling Mutual?

### 2014-08-27 mutual-modeling.tex
Completed many of the missing ref, switched to Harvard-style citations

@@

@@

\begin{document}
\maketitle

@@

From its very beginning, CSCL research has been following Roschelle \& Teasley's
(1986) suggestion that collaborative learning has something to do with the
process of constructing and maintaining a shared understanding of the task at
hand. Building a shared/mutual understanding refers to the upper class of
collaborative learning situations, those in which students should build upon
each other's understanding to refine their own understanding. What is expected
to produce learning is not the mere fact that two students build the same
understanding but the cognitive effort they have to engage to build this shared
understanding (Schwartz, 1995). This effort can be observed by the frequency of
rich interactions, \ie interactions whose occurrence has been related to
learning: (self-) explanations in cognitive science (Chi ; Webb), conflict
resolution in socio-cognitive theories (Doise), mutual regulation (Blaye,) in a
Vygostkian perspective, etc. The construction of a shared understanding has been
investigated for several years in psycholinguistics, under the  notion of
“grounding” (Clark \&Wilkes-Gibbs, 1986). However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned. The monitoring
and repair of mis-understanding explains for instance referential failures in
short dialogue episodes but does hardly predict conceptual change over longer
sessions (Dillenbourg \& Traum, 2006). The cumulative effect of grounding
episodes can probably be better understood from a socio-cultural perspective:
"collaborative learning is associated with the increased cognitive-interactional
effort involved in the transition from learning to understand each other to
learning to understand the meanings of the semiotic tools that constitute the
mediators of interpersonal interaction" (Baker, Hansen, Joiner \& Traum, 1999).
Several scholars suggest that CSCL research should go deeper in the
understanding of how partners engage into shared meaning making (Stahl, 2006) or
'intersubjective' meaning making (Suthers (2006).

Paradoxically, while Clark's theory is somewhat too linguistic from a conceptual
change viewpoint, it is criticized at the same time as being too cognitivist by
some psycholinguists, \ie as overestimating the amount shared knowledge and
mutual representations actually necessary to conduct a dialogue. The fundamental
issue, as old as philosophy, is the degree of coupling between the different
levels of dialogue, mostly between the lexical / syntactical level and the
deeper semantic levels. Pickering \& Garrod (2004) argue that the mutual
understanding starts mostly with a 'local alignment' at the level of the
linguistic representations, due to priming mechanisms, and that this superficial
alignment may – in some cases- lead to a 'global alignment' of the semantic
level.  For these authors, the convergence in dialogue, and even the repair of
some misunderstandings, is explained by this mimetic behavior more than by a
monitoring of each other's knowledge: "…interlocutors do not need to monitor and
develop full common ground as a regular, constant part of routine conversation,
as it would be unnecessary and far too costly. Establishment of full common
ground is, we argue, a specialized and non-automatic process that is used
primarily in times of difficulty (when radical misalignment becomes apparent)."
(p. 179). This view is actually not incompatible with Clark's grounding
criterion (REF): the degree of shared understanding that peers need to reach
depends upon the task they perform. For instance, a dialogue between two
surgeons might rely on superficial alignment if they talk about their friends
but has to guarantee accurate common grounds when talking about which
intervention will be conducted in which way on which patient.  Some CSCL
scholars stressed the 'illusion of shared understanding' (JACCO?), \ie the
potential decoupling between linguistic alignment and actual shared meanings.

This interesting cognitive science debate mostly occurred outside the field of
learning. In education, the question is to relate these mechanisms to learning
outcomes. Is linguistic alignment sufficient to trigger conceptual change? Does
negotiation of meaning only occurs when partners monitor and diagnose each
other's knowledge. If the ratio between shallow alignment and deep grounding
depends upon the task, and if deep grounding is a condition for learning, then
the pedagogical challenge is to design tasks that require deep grounding. Most
empirical studies on grounding and alignment are conducted with tasks, which
despite being qualified as 'ecologically valid' by their authors, are mere
referencing tasks such as asking the way to the train station or helping the
peer to choose a picture among many. In this contribution, we explore several
richer tasks such as arguing on a sensitive issue or building a concept map.

Deep grounding or shared meaning making requires some cognitive load. For Clark
\& Wilkes-Gibbs (1986), what is important is not the individual effort made by
the receiver of a communicative act, but the overall least collaborative effort.
The cost of producing a perfect utterance may be higher than the cost of
repairing the problems that may arise through misunderstandings. For instance,
subjects are less careful about adapting their utterances to their partner when
they know they can provide feedback on his/her understanding (Schober, 1993). We
introduced the notion of ‘optimal collaborative effort' (Dillenbourg et al,
1996) to stress that misunderstanding should not be viewed as something to be
avoided (if this was possible), but as an opportunity to engage into
verbalization, explanation, negotiation, and so forth. This issue is related to
the global argument regarding cognitive load in learning activities, especially
in discovery learning environments: there is no learning without some cognitive
load, but overload may hinder learning (Paas, Renkl \& Sweller, 2003). In the
context of collaborative learning, we understand the cognitive load induced by
mutual modeling as part of Schwartz's (1995) notion of effort towards a shared
understanding. For instance, CSCL researchers expanded the use of 'collaboration
scripts' (Aronson,?). A script is a pedagogical method that collaborative
learning activities in order to foster the emergence of productive interactions
such as argumentation, explanation or conflict. Conflict-resolution scripts such
as the ArgueGraph (Dillenbourg \& Hong, 2008) forms pair of students with
opposite opinions, which increases the difficulty of consensus building,
requiring more justifications, more negotiation, more load. Similarly, JIGSAW
scripts provide peers with different but complementary knowledge for augmenting
(reasonably) the effort group members have to engage to reach a shared solution.

%\paragraph{Theory of Mind}
%
%Theory of Mind (originally defined in~\cite{premack1978does}) is the cognitive
%ability that a subject possesses to represent the mental state of another
%agent, possibly including knowledge that contradicts the subject's own model: for
%example, a book can be at the same time \emph{visible} for myself, and \emph{not
%visible} for you.
%
%Children develop this skill, which is essential to understand others'
%perspectives during interactions, around the age of three. It supposes the
%ability to build, store and retrieve separate models of the knowledge of the
%interactors.
%
%One classical application of this cognitive skill is the so-called
%\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
%experiment)~\cite{Leslie2000}: a child is asked to watch a scene where two
%people, A and B, manipulate objects. Then A leaves and B hides away one
%object. When A comes back, we ask the child ``where do you think A will
%look for the object?''. Before acquiring a theory of mind, children are not
%able to separate their own (true) model of the world (where they know that
%the object was hidden) from the model of A, which contains \emph{false
%beliefs} on the world (A still thinks the object is at its original
%position since he did not see B hiding it).

@@

The content of mutual models ranges from 'dispositional' versus 'situational'
aspects. The 'dispositional' aspects refer to A's representation of B's long
term knowledge, skills or traits. It is thus closely related to the notion of
transactive memory~\cite{wegner1987transactive, moreland1999transactive}.
'Situational' aspects refer to A's representation of B's knowledge, behavior or
intentions specifically activated in the situation in which A and B are
collaborating, some of them being valid for 2 seconds, other ones for 2 hours.
Here are examples of fragments that constitute A's model of B regarding to
aspects X, noted $Model(A,B,X)$:

@@

A partner model is probably not a "box", \ie not a monolithical representation
but rather a mosaic of information fragments about the partner, with various
grain sizes and various life cycles. This mosaic is elaborated through various
mechanisms, first for building an initial model of the partner and then for
updating this model.  As two students meet for the first time, mutual models are
initialized by the assumptions they make upon each other from cues such as
his/her membership to large categories (age, culture, profession, ...) include
stereotypes (sportsmen, junkie, business women, Swiss,...) as well as physical
appearance. Scholars studied how initial modeling impacts communication. In
their experiments on initial mutual modelling, Slugoski, Lalljee, Lamb \&
Ginsburg~\cite{slugoski1993attribution} pretended to their subjects that their
(fake) partner had or had not received the same information. They observed that
the subjects adapted their dialogue by focusing the explanation on the items
that (s)he was supposed to ignore. Brennan (1991) \fixme{ref} showed that the subjects used
different initial strategies in forming queries depending on who they were told
their partner was.

Initial common grounds are also initiated by co-presence: they include events to
which A and B attended together~\cite{clark2002definite} in the physical space
or in our cultural space (\eg ‘09-11'). Co-presence means that we can refer to
a shared objects and event,-s it does of course not imply that we give them the
same meaning. Namely, a shared screen does not mean a shared
understanding~\cite{dillenbourg2006sharing}.

@@

Mutual modeling does not occur in a vacuum but it is highly contextualized. REF
reviews the features of the collaborative situation, namely the media
(cotemporality… ), may facilitate or hamper mutual modeling. Hutchins (SILENCE)
reported a study in which a short silence between two pilots was perfectly
interpreted because it occurred in a highly constrained communication context.
Some environments are more productive than others in helping peers to detect
their misunderstandings. Roschelle~\cite{roschelle1995construction} reformulate
the design of CSCL interfaces as providing ways for peer to detect and repairs
their misunderstanding. In CSCW, researchers have explored various so-called
"awareness tools", \ie functionalities that inform A is about B's actions that A
could not directly perceived because B was working on a different subset of the
virtual space. Different awareness tools will be addressed in this contribution
since they allowed us to manipulate experimentally the mutual modelling activity.

%\subsection{State of the Art}
%
%\paragraph{Building a model of a human during the interaction} Perspective
%Taking is a human ability which allows one to put him/herself in another
%person's point of view. Studied in psychology
%literature~\cite{Flavell1992,Tversky1999}, this ability is crucial when
%interacting with people by allowing one to reason on others' understanding of
%the world in terms of visual perception, spatial descriptions, affordances and
%beliefs, etc.  In the last years these notions have been gradually employed in
%HRI.~\cite{Breazeal2006} presents a learning algorithm that takes into account
%information about a teacher's visual perspective in order to learn a task.
%~\cite{Johnson2005} apply visual perspective taking for action recognition
%between two robots.~\cite{Trafton2005} use both visual and spatial perspective
%taking to find out the referent indicated by a human partner.
%
%The cognitive model that the robot builds for the agent it interacts with is
%still simple and mostly focused on geometric features (who sees what? What are
%our relative positions? etc.). Extending this knowledge with more subtle
%perceptions (emotional state for instance) remains to be done.
%
%\paragraph{Joint action}
%Several theories dealing with collaboration~\cite{Cohen1991,Grosz1996,Clark1996}
%emphasize that collaborative tasks have specific requirements compared to
%individual ones, \eg, since the robot and the person share a common goal, they
%have to agree on the manner to realize it, they must show their commitment to
%the goal during execution, etc. Several robotic systems have already been built
%based on these theories~\cite{Rich1997,Sidner2005,Breazeal2003} and they all
%have shown benefits of this approach. They have also shown how difficult it is
%to manage turn-taking between communication partners and to interleave task
%realization and communication in a generic way. Finally, today only few
%systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
%levels.
%
\section{A model for mutual modelling evaluation}

@@

Based on these 3 parameters, the probability of making an action $\alpha_{t_1}$ about
content X at time $t_1$ during task T in order to increase \Model{A}{B}{X}
at time $t_2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
costs (3)~\cite{traum1996miscommunication}:

@@

Along the same reasoning, the initial degree of common grounds \inigrounding
should be rather low (and hence the difference between initial and required
degrees rather high) in order to make mutual modelling effort more observable.
While we can exactly control the actual initial knowledge available to the
robot, estimating \inigrounding still requires special care during the design of
the experiments: as we have previously shown~\cite{lemaignan2014dynamics,
lemaignan2014cognitive} and because the interaction takes place between a human and
a robot, the initial human model of the robot can widely vary depending on
several factors like the user familiarity with technology, the effect of
novelty, the user expectations or the anthropomorphic features of the robot.

@@

We studied the impact of an awareness tool on group performance and mutual
modeling~\cite{nova2007collaboration}. The availability
of an awareness tool was our independent variable. In previous studies, we
replayed a video of the game to subjects who surprised us by their ability to
remember former states of their mutual model: ``I did that because I thought that
you would do that''. Hence, this experiment focused the representation of each
other's action plans. During the game, we asked them to anticipate the next
action of their partner as well as to announce their own action. It is clear
asking them a question about their partner increases their attention for the
rest of the game.

@@

\paragraph{Symmetry question} We computed intra-class correlation as described
by~\cite{kenny1998data} from the answers to the cross-questionnaires.
Considering all pairs in both conditions, we found a positive and significant
correlation ($r = .38, p < .05$) between \gModel{A}{B} and \gModel{B}{A}.
Interestingly, this was higher in the control group ($r = 0.44$) than in the
experimental group ($r = 0.24$). Actually, $\Delta(\gMdeg{A}{B},\gMdeg{B}{A})$,
\ie the absolute differences between the models accuracy, was not significantly
different with or without the awareness tool ($F [1,13]= 0.144, p > 0.5$). This
result could be explained by the fact that the players without awareness tools
communicated more.

@@

This study concerns a collaborative game that occurred in the physical
space~\cite{nova2006underwhelming}. We studied whether players build an accurate
model of the path followed by their partners because this path reflected their
problem solving strategy. We used an objective measure of \gModel{A}{B}: the
distance between where A believes B has been walking and where B actually went.
The main hypothesis concerned the effect of awareness tools on group performance
and on \gModel{A}{B}.

@@

\paragraph{Study-specific questions} There was no significant difference
regarding the task performance. However, our surprise was that the absence of
the awareness tool was related to a higher \gModel{A}{B}: players better
remembered their partners' path if they did not see their position permanently!
We will not enter into the details of these results but, to be short, teams
without awareness tool made more annotations on the map while permanent MLA has
an underwhelming effect~\cite{nova2005location}.

@@

This positive result regarding to the symmetry question is even more surprising
than the previous one: despite the high heterogeneity of spatial skills among
adults~\cite{liben1981spatial}, a significant correlation could indicate that
mutual modeling relies more on group processes than on individual processes. The
significant results on the triangle questions probably go in the same direction:
either the modeler or the modeled both contribute to accurate modeling, or that
the quality of teamwork has a global effect. In this experiment, since team
members did not interact intensively during the task, the factor that
facilitates accurate mutual modeling is probably less the quality of verbal
interactions that the group strategy: a clear strategy facilitates memorizing
one's partner path.

@@

\paragraph{Study-specific questions} Our hypothesis about media richness is
rejected: the average \Model{A}{B}{emotions} was .62 ($SD = .28$) with the
audio+video condition and .71 ($SD = .19$) in the audio condition ($F=3.4, p:
0.065$). Our interpretation is that viewing what one's partner sees (shared
graphical editor) was, in this task, more important than seeing each
other~\cite{gaver1993one}\fixme{check ref with Pierre}\fixme{other cite:
Anderson et al., 2000}.

@@

This study investigated the effect of a collaboration script on collaborative
learning~\cite{molinari2008effects}. The script chosen is a JIGSAW: two students
receive different but complementary subsets of the knowledge (texts) which have
to be integrated to build a shared concept map.  This script increases the
cognitive effort to build the map, not only to recon ciliate the viewpoints of
each team members but, before that, to find out what the other knows.

@@

The peers were located into two different rooms. A complete description of the
study is provided in~\cite{sangin2008learners}. The experiment lasted 90 minutes.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrv}
\bibliography{biblio}

>>>>

% UTF8 support
\usepackage[utf8x]{inputenc}

@@

\usepackage[round]{natbib}

@@

\author{
    Pierre Dillenbourg, Séverin Lemaignan \\
    \small Computer-Human Interaction in Learning and Instruction (CHILI) \\
    \small Ecole Polytechnique Fédérale de Lausanne (EPFL) \\
    \small CH-1015 Lausanne, Switzerland \\
    {\tt firstname.lastname@epfl.ch}
    \and
    Mirweis Sangin \\
    \and
    Nicolas Nova \\
    \small Haute Ecole Arts \& Design de Genève\\
    {\tt nicolas.nova@hesge.ch}
    \and
    René Glaus
    \and
    Gaëlle Molinari \\
    \small Fernuni Schweiz \\
    {\tt gaelle.molinari@fernuni.ch}
}
\begin{document}
\maketitle

@@

From its very beginning, CSCL research has been following
\citet{roschelle1995construction} suggestion that collaborative learning has
something to do with the process of constructing and maintaining a shared
understanding of the task at hand. Building a shared/mutual understanding refers
to the upper class of collaborative learning situations, those in which students
should build upon each other's understanding to refine their own understanding.
What is expected to produce learning is not the mere fact that two students
build the same understanding but the cognitive effort they have to engage to
build this shared understanding~\citep{schwartz1995emergence}. This effort can
be observed by the frequency of rich interactions, \ie interactions whose
occurrence has been related to learning: (self-) explanations in cognitive
science (Chi ; Webb), conflict resolution in socio-cognitive theories (Doise),
mutual regulation (Blaye,) in a Vygostkian perspective, etc. The construction of
a shared understanding has been investigated for several years in
psycholinguistics, under the  notion of
``grounding''~\citep{clark1986referring}.  However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned. The monitoring
and repair of mis-understanding explains for instance referential failures in
short dialogue episodes but does hardly predict conceptual change over longer
sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
episodes can probably be better understood from a socio-cultural perspective:
``collaborative learning is associated with the increased
cognitive-interactional effort involved in the transition from learning to
understand each other to learning to understand the meanings of the semiotic
tools that constitute the mediators of interpersonal
interaction''~\citep{baker1999role}.  Several scholars suggest that CSCL
research should go deeper in the understanding of how partners engage into
shared meaning making~\citep{stahl2007meaning} or \emph{intersubjective} meaning
making~\citep{suthers2006technology}.

Paradoxically, while Clark's theory is somewhat too linguistic from a conceptual
change viewpoint, it is criticized at the same time as being too cognitivist by
some psycholinguists, \ie as overestimating the amount shared knowledge and
mutual representations actually necessary to conduct a dialogue. The fundamental
issue, as old as philosophy, is the degree of coupling between the different
levels of dialogue, mostly between the lexical / syntactical level and the
deeper semantic levels. \citet{pickering2006alignment} argue that the mutual
understanding starts mostly with a \emph{local alignment} at the level of the
linguistic representations, due to priming mechanisms, and that this superficial
alignment may -- in some cases -- lead to a \emph{global alignment} of the
semantic level.  For these authors, the convergence in dialogue, and even the
repair of some misunderstandings, is explained by this mimetic behavior more
than by a monitoring of each other's knowledge: ``...interlocutors do not need
to monitor and develop full common ground as a regular, constant part of routine
conversation, as it would be unnecessary and far too costly. Establishment of
full common ground is, we argue, a specialized and non-automatic process that is
used primarily in times of difficulty (when radical misalignment becomes
apparent).'' (p. 179). This view is actually not incompatible with Clark's
grounding criterion \fixme{(REF)}: the degree of shared understanding that peers
need to reach depends upon the task they perform. For instance, a dialogue
between two surgeons might rely on superficial alignment if they talk about
their friends but has to guarantee accurate common grounds when talking about
which intervention will be conducted in which way on which patient.  Some CSCL
scholars stressed the ``illusion of shared understanding'' \fixme{REF:(JACCO?)},
\ie the potential decoupling between linguistic alignment and actual shared
meanings.

This interesting cognitive science debate mostly occurred outside the field of
learning. In education, the question is to relate these mechanisms to learning
outcomes. Is linguistic alignment sufficient to trigger conceptual change? Does
negotiation of meaning only occurs when partners monitor and diagnose each
other's knowledge. If the ratio between shallow alignment and deep grounding
depends upon the task, and if deep grounding is a condition for learning, then
the pedagogical challenge is to design tasks that require deep grounding. Most
empirical studies on grounding and alignment are conducted with tasks, which
despite being qualified as ``ecologically valid'' by their authors, are mere
referencing tasks such as asking the way to the train station or helping the
peer to choose a picture among many. In this contribution, we explore several
richer tasks such as arguing on a sensitive issue or building a concept map.

Deep grounding or shared meaning making requires some cognitive load. For
\citet{clark1986referring}, what is important is not the individual effort made
by the receiver of a communicative act, but the overall least collaborative
effort.  The cost of producing a perfect utterance may be higher than the cost
of repairing the problems that may arise through misunderstandings. For
instance, subjects are less careful about adapting their utterances to their
partner when they know they can provide feedback on his/her
understanding~\citep{schober1993spatial}. We introduced the notion of
\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
that misunderstanding should not be viewed as something to be avoided (if this
was possible), but as an opportunity to engage into verbalization, explanation,
negotiation, and so forth. This issue is related to the global argument
regarding cognitive load in learning activities, especially in discovery
learning environments: there is no learning without some cognitive load, but
overload may hinder learning~\citep{paas2003cognitive}. In the context of
collaborative learning, we understand the cognitive load induced by mutual
modeling as part of~\citet{schwartz1995emergence} notion of effort towards a
shared understanding. For instance, CSCL researchers expanded the use of
\emph{collaboration scripts} \fixme{REF (Aronson,?)}. A script is a pedagogical
method that collaborative learning activities in order to foster the emergence
of productive interactions such as argumentation, explanation or conflict.
Conflict-resolution scripts such as the {\sc
ArgueGraph}~\citep{dillenbourg2008mechanics} forms pair of students with
opposite opinions, which increases the difficulty of consensus building,
requiring more justifications, more negotiation, more load. Similarly, JIGSAW
scripts provide peers with different but complementary knowledge for augmenting
(reasonably) the effort group members have to engage to reach a shared solution.

%\paragraph{Theory of Mind}
%
%Theory of Mind (originally defined in~\citep{premack1978does}) is the cognitive
%ability that a subject possesses to represent the mental state of another
%agent, possibly including knowledge that contradicts the subject's own model: for
%example, a book can be at the same time \emph{visible} for myself, and \emph{not
%visible} for you.
%
%Children develop this skill, which is essential to understand others'
%perspectives during interactions, around the age of three. It supposes the
%ability to build, store and retrieve separate models of the knowledge of the
%interactors.
%
%One classical application of this cognitive skill is the so-called
%\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
%experiment)~\citep{Leslie2000}: a child is asked to watch a scene where two
%people, A and B, manipulate objects. Then A leaves and B hides away one
%object. When A comes back, we ask the child ``where do you think A will
%look for the object?''. Before acquiring a theory of mind, children are not
%able to separate their own (true) model of the world (where they know that
%the object was hidden) from the model of A, which contains \emph{false
%beliefs} on the world (A still thinks the object is at its original
%position since he did not see B hiding it).

@@

The content of mutual models ranges from 'dispositional' versus 'situational'
aspects. The 'dispositional' aspects refer to A's representation of B's long
term knowledge, skills or traits. It is thus closely related to the notion of
transactive memory~\citep{wegner1987transactive, moreland1999transactive}.
'Situational' aspects refer to A's representation of B's knowledge, behavior or
intentions specifically activated in the situation in which A and B are
collaborating, some of them being valid for 2 seconds, other ones for 2 hours.
Here are examples of fragments that constitute A's model of B regarding to
aspects X, noted $Model(A,B,X)$:

@@

A partner model is probably not a "box", \ie not a monolithical representation
but rather a mosaic of information fragments about the partner, with various
grain sizes and various life cycles. This mosaic is elaborated through various
mechanisms, first for building an initial model of the partner and then for
updating this model.  As two students meet for the first time, mutual models are
initialized by the assumptions they make upon each other from cues such as
his/her membership to large categories (age, culture, profession, ...) include
stereotypes (sportsmen, junkie, business women, Swiss,...) as well as physical
appearance. Scholars studied how initial modeling impacts communication. In
their experiments on initial mutual modelling, Slugoski, Lalljee, Lamb \&
Ginsburg~\citep{slugoski1993attribution} pretended to their subjects that their
(fake) partner had or had not received the same information. They observed that
the subjects adapted their dialogue by focusing the explanation on the items
that (s)he was supposed to ignore. Brennan (1991) \fixme{ref} showed that the subjects used
different initial strategies in forming queries depending on who they were told
their partner was.

Initial common grounds are also initiated by co-presence: they include events to
which A and B attended together~\citep{clark2002definite} in the physical space
or in our cultural space (\eg ‘09-11'). Co-presence means that we can refer to
a shared objects and event,-s it does of course not imply that we give them the
same meaning. Namely, a shared screen does not mean a shared
understanding~\citep{dillenbourg2006sharing}.

@@

Mutual modeling does not occur in a vacuum but it is highly contextualized. REF
reviews the features of the collaborative situation, namely the media
(cotemporality… ), may facilitate or hamper mutual modeling. Hutchins (SILENCE)
reported a study in which a short silence between two pilots was perfectly
interpreted because it occurred in a highly constrained communication context.
Some environments are more productive than others in helping peers to detect
their misunderstandings. Roschelle~\citep{roschelle1995construction} reformulate
the design of CSCL interfaces as providing ways for peer to detect and repairs
their misunderstanding. In CSCW, researchers have explored various so-called
"awareness tools", \ie functionalities that inform A is about B's actions that A
could not directly perceived because B was working on a different subset of the
virtual space. Different awareness tools will be addressed in this contribution
since they allowed us to manipulate experimentally the mutual modelling activity.

%\subsection{State of the Art}
%
%\paragraph{Building a model of a human during the interaction} Perspective
%Taking is a human ability which allows one to put him/herself in another
%person's point of view. Studied in psychology
%literature~\citep{Flavell1992,Tversky1999}, this ability is crucial when
%interacting with people by allowing one to reason on others' understanding of
%the world in terms of visual perception, spatial descriptions, affordances and
%beliefs, etc.  In the last years these notions have been gradually employed in
%HRI.~\citep{Breazeal2006} presents a learning algorithm that takes into account
%information about a teacher's visual perspective in order to learn a task.
%~\citep{Johnson2005} apply visual perspective taking for action recognition
%between two robots.~\citep{Trafton2005} use both visual and spatial perspective
%taking to find out the referent indicated by a human partner.
%
%The cognitive model that the robot builds for the agent it interacts with is
%still simple and mostly focused on geometric features (who sees what? What are
%our relative positions? etc.). Extending this knowledge with more subtle
%perceptions (emotional state for instance) remains to be done.
%
%\paragraph{Joint action}
%Several theories dealing with collaboration~\citep{Cohen1991,Grosz1996,Clark1996}
%emphasize that collaborative tasks have specific requirements compared to
%individual ones, \eg, since the robot and the person share a common goal, they
%have to agree on the manner to realize it, they must show their commitment to
%the goal during execution, etc. Several robotic systems have already been built
%based on these theories~\citep{Rich1997,Sidner2005,Breazeal2003} and they all
%have shown benefits of this approach. They have also shown how difficult it is
%to manage turn-taking between communication partners and to interleave task
%realization and communication in a generic way. Finally, today only few
%systems~\citep{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
%levels.
%
\section{A model for mutual modelling evaluation}

@@

Based on these 3 parameters, the probability of making an action $\alpha_{t_1}$ about
content X at time $t_1$ during task T in order to increase \Model{A}{B}{X}
at time $t_2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
costs (3)~\citep{traum1996miscommunication}:

@@

Along the same reasoning, the initial degree of common grounds \inigrounding
should be rather low (and hence the difference between initial and required
degrees rather high) in order to make mutual modelling effort more observable.
While we can exactly control the actual initial knowledge available to the
robot, estimating \inigrounding still requires special care during the design of
the experiments: as we have previously shown~\citep{lemaignan2014dynamics,
lemaignan2014cognitive} and because the interaction takes place between a human and
a robot, the initial human model of the robot can widely vary depending on
several factors like the user familiarity with technology, the effect of
novelty, the user expectations or the anthropomorphic features of the robot.

@@

We studied the impact of an awareness tool on group performance and mutual
modeling~\citep{nova2007collaboration}. The availability
of an awareness tool was our independent variable. In previous studies, we
replayed a video of the game to subjects who surprised us by their ability to
remember former states of their mutual model: ``I did that because I thought that
you would do that''. Hence, this experiment focused the representation of each
other's action plans. During the game, we asked them to anticipate the next
action of their partner as well as to announce their own action. It is clear
asking them a question about their partner increases their attention for the
rest of the game.

@@

\paragraph{Symmetry question} We computed intra-class correlation as described
by~\citep{kenny1998data} from the answers to the cross-questionnaires.
Considering all pairs in both conditions, we found a positive and significant
correlation ($r = .38, p < .05$) between \gModel{A}{B} and \gModel{B}{A}.
Interestingly, this was higher in the control group ($r = 0.44$) than in the
experimental group ($r = 0.24$). Actually, $\Delta(\gMdeg{A}{B},\gMdeg{B}{A})$,
\ie the absolute differences between the models accuracy, was not significantly
different with or without the awareness tool ($F [1,13]= 0.144, p > 0.5$). This
result could be explained by the fact that the players without awareness tools
communicated more.

@@

This study concerns a collaborative game that occurred in the physical
space~\citep{nova2006underwhelming}. We studied whether players build an accurate
model of the path followed by their partners because this path reflected their
problem solving strategy. We used an objective measure of \gModel{A}{B}: the
distance between where A believes B has been walking and where B actually went.
The main hypothesis concerned the effect of awareness tools on group performance
and on \gModel{A}{B}.

@@

\paragraph{Study-specific questions} There was no significant difference
regarding the task performance. However, our surprise was that the absence of
the awareness tool was related to a higher \gModel{A}{B}: players better
remembered their partners' path if they did not see their position permanently!
We will not enter into the details of these results but, to be short, teams
without awareness tool made more annotations on the map while permanent MLA has
an underwhelming effect~\citep{nova2005location}.

@@

This positive result regarding to the symmetry question is even more surprising
than the previous one: despite the high heterogeneity of spatial skills among
adults~\citep{liben1981spatial}, a significant correlation could indicate that
mutual modeling relies more on group processes than on individual processes. The
significant results on the triangle questions probably go in the same direction:
either the modeler or the modeled both contribute to accurate modeling, or that
the quality of teamwork has a global effect. In this experiment, since team
members did not interact intensively during the task, the factor that
facilitates accurate mutual modeling is probably less the quality of verbal
interactions that the group strategy: a clear strategy facilitates memorizing
one's partner path.

@@

\paragraph{Study-specific questions} Our hypothesis about media richness is
rejected: the average \Model{A}{B}{emotions} was .62 ($SD = .28$) with the
audio+video condition and .71 ($SD = .19$) in the audio condition ($F=3.4, p:
0.065$). Our interpretation is that viewing what one's partner sees (shared
graphical editor) was, in this task, more important than seeing each
other~\citep{gaver1993one}\fixme{check ref with Pierre}\fixme{other cite:
Anderson et al., 2000}.

@@

This study investigated the effect of a collaboration script on collaborative
learning~\citep{molinari2008effects}. The script chosen is a JIGSAW: two students
receive different but complementary subsets of the knowledge (texts) which have
to be integrated to build a shared concept map.  This script increases the
cognitive effort to build the map, not only to recon ciliate the viewpoints of
each team members but, before that, to find out what the other knows.

@@

The peers were located into two different rooms. A complete description of the
study is provided in~\citep{sangin2008learners}. The experiment lasted 90 minutes.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plainnat}
\bibliography{biblio}
